---
title: "Ocean Acidification"
author: "Carlos J. Arroyo Ruiz"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
Main Question: Can we quantify ocean acidification in the subtropical north Atlantic Ocean?
1) Is the surface ocean pCO2 increasing?
2) Is the surface ocean pH decreasing?
3) Is the surface ocean seawater saturation state with respect to aragonite decreasing?

Assigment Format
Intro > Methods > Results > Discussion mini-lab report

### Load required libraries
```{r}
library(tidyverse)
library(seacarb)
library(performance)
```

###Import Data from BATS
```{r}
bats_bottle <- read_delim("bats_bottle.txt", delim = "\t", escape_double = FALSE, 
                          col_names = FALSE, trim_ws = TRUE, skip = 60)
View(bats_bottle)

colnames(bats_bottle) <- colnames(read_csv("bats_bottle.txt", skip = 59))
#Check that I have done this correctly!
View(bats_bottle)
#I have a data frame of BATS data with column names for variables
```

Variable Units
-yyyymmdd = Year Month Day   
decy   = Decimal Year     
time   = Time (hhmm)      
latN   = Latitude (Deg N) 
lonW   = Longitude (Deg W)
Depth  = Depth (m)                  
Temp   = Temperature ITS-90 (C)    
CTD_S  = CTD Salinity (PSS-78)      
Sal1   = Salinity-1 (PSS-78)        
Sig-th = Sigma-Theta (kg/m^3)       
O2(1)  = Oxygen-1 (umol/kg)          
OxFixT = Oxygen Fix Temp (C)        
Anom1  = Oxy Anomaly-1 (umol/kg)    
CO2    = dissolved inorganic carbon (umol/kg)              
Alk    = Alkalinity (uequiv)        
NO31   = Nitrate+Nitrite-1 (umol/kg)
NO21   = Nitrite-1 (umol/kg)        
PO41   = Phosphate-1 (umol/kg)      
Si1    = Silicate-1 (umol/kg)       
POC    = POC (ug/kg)                
PON    = PON (ug/kg)                
TOC    = TOC (umol/kg)                
TN     = TN (umol/kg)  
Bact   = Bacteria enumeration (cells*10^8/kg)   
POP    = POP (umol/kg)
TDP    = Total dissolved Phosphorus (nmol/kg)
SRP    = Low-level phosphorus (nmol/kg)
BSi    = Particulate biogenic silica (umol/kg)
LSi    = Particulate lithogenic silica  (umol/kg)
Pro    = Prochlorococcus (cells/ml)
Syn    = Synechococcus (cells/ml)
Piceu  = Picoeukaryotes (cells/ml)
Naneu  = Nanoeukaryotes (cells/ml)

### Calculate CO2 chemistry parameters

```{r}
?carb(seacarb)
?carb
#carb(flag, var1, var2, S=35, T=25, Patm=1, P=0, Pt=0, Sit=0,
        #k1k2="x", kf="x", ks="d", pHscale="T", b="u74", gas="potential", 
        #warn="y", eos="eos80", long=1.e20, lat=1.e20)
#We have TA, DIC, S, T, Pt, Sit, but we don't have pressure.

#First we need to calculate pressure (dbar) using TEOS-10:
?gsw
#p= gsw_p_from_z(z,lat)
bats_co2=bats_bottle %>%
  mutate(P_dbar=gsw_p_from_z(Depth*-1,latN))
View(bats_co2)

#We now have all of the variables that we neeed to calculate the surface seawater chemistry at the BATS station, but we need to be very careful about our units!

#We now have TA, DIC, S, Pt, Sit, and Pressure
#What are  the units of these what does CO2SYS need?
#Ta is in uequiv (umol/kg) and we need mol/kg
#S is in PSS and we will use EOS80
#T is in degrees C and we need degrees C
#Pt is in umol/kg and we need mol/kg
#SIT is in umol/kg and we need mol/kg
#P_dbar is in dbar and we need bar

#We will need to convert units scaling when using CO2SYS

#flag = 15 ALK and DIC given
?carb
bats_co2sys=
  bats_co2 %>% 
  filter(Alk!=-999,CO2!=-999,Sal1!=-999,Temp!=-999,
         P_dbar!=-999,PO41!=-999,Si1!=-999,
         lonW!=-999,latN!=-999)%>% 
  rename(DIC=CO2) %>% 
  mutate(carb(flag=15, var1=Alk*10^-6, var2=DIC*10^-6, 
              S=Sal1, T=Temp, Patm=1, P=P_dbar*10^-1, 
              Pt=PO41*10^-6, Sit=Si1*10^-6,
              k1k2="l", kf="pf", ks="d", pHscale="T", 
              b="u74", gas="potential", 
              warn="y", eos="eos80",long=360-lonW, lat=latN))

#filter for only the surface ocean
bats_co2sys_surf = bats_co2sys %>% 
  filter(Depth<100) %>% #we are selecting only the upper 100m
  filter(pCO2insitu>100) #BATS told me this data point is bad.

#1) Is surface ocean pCO2 increasing?

bats_co2sys_surf %>% 
  ggplot(mapping=aes(x=decy,y=pCO2insitu))+
  geom_point()+
  geom_smooth()

check=bats_co2sys_surf %>% 
  filter(pCO2<200)

pco2_model = lm(pCO2insitu~decy,data=bats_co2sys_surf)
summary(pco2_model)

#For Homework:
#Make decy vs PCO2 plot pretty

bats_co2sys_surf %>% 
  ggplot(mapping=aes(x=decy,y=pCO2insitu))+
  geom_line()+
  geom_smooth()+
  scale_y_continuous()+
  scale_x_continuous()+
  xlab(expression(paste("Year")))+
  ylab(expression(paste("Partial Pressure CO"[2], "(In situ)")))+
  theme_classic()

#Could include methods text for CO2SYS in lab report

#For Thursday:
#How do we check our model performance?
#How do we plot our model predictions?

#2) IS surface ocean seawater pH decreasing?
#3) Is surface ocean seawater saturation state with respect to aragonite decreasing?
#These two questions for thursday

################################################################################################
#How do we check our model performance?

#Checking model performance
#install.packages("performance")
library(performance)
library(see)

check_model(pco2_model)

summary(pco2_model)
anova(pco2_model) #p-value is the same as summary

#How do plot our model prediction?

#The tidy way is with diplyr 

bats_co2sys_surf_pred=
bats_co2sys_surf %>%
  mutate(predict(pco2_model,interval='confidence', level=0.95))
#the base r way with cbind an do not need to rename outputs

bats_co2sys_surf_pred=
  cbind(bats_co2sys_surf, predict(pco2_model, interval='confidence', level=0.95))

#predict(pco2_model,as.data.frame(decy=seq(1990,2020,0.1)), interval='confidence',level=0.95)

bats_co2sys_surf_pred %>% 
  ggplot()+
  geom_point(mapping=aes(x=decy,y=pCO2insitu))+
  geom_line(aes(decy,fit),color="purple")+
  geom_ribbon(aes(x=decy,ymin=lwr,ymax=upr),alpha=0.2,fill="purple")+
  theme_classic()

#1) Is surface ocean pCO2 increasing?
#There is a seasonal cycle in the surface ocean pCO2 at BATS with a higher pCO2 observed in late summer to early fall and lower pCO2 observed winter and early spring. There is a consistent, detectable (ANOVA,p<0.001) annual increase in pCO2 by 1.85+-0.07 uatm/year.
#Also show plot with model. The figure caption describes plot (points=data, line=model, shaded region = 95% confidence interval.)

#Complete questions 1,2,3 in finishes R markdown document. You can also include a map if you'd like to practice google maps!

###################################################
####How can we improve our pCO2 predictions?
#What might be some valuable predictors of seawater pCO2?
# Temperature - measure of atmospheric pCO2, but alos partial pressure of CO2 in seawater.
# Seasons - months (winter vs summer)
# Years - annual increases in atmospheric CO2
# Dissolved Oxygen - photosynthesis and respiration impact both DO and CO2.
# Salinity - impacts solubility
# Nutrients - also related through photosynthesis

#Create a year and month column with BATS data
bats_co2sys_surf$year=
  as.numeric(substr(bats_co2sys_surf$yyyymmdd,1,4))
bats_co2sys_surf$month=
  as.numeric(substr(bats_co2sys_surf$yyyymmdd,5,6))

m1=lm(pCO2insitu~decy,data=bats_co2sys_surf)
m2=lm(pCO2insitu~year+month,data=bats_co2sys_surf)
m3=lm(pCO2insitu~year+month+Temp,data=bats_co2sys_surf)

summary(m1) #r^2=0.3
summary(m2) #r^2=0.4
summary(m3) #r^2=0.7
AIC(m1,m2,m3) #Use AIC to select models (lower AIC = better model)

#For model development, I am interested in
#year, month, temp, sal, o2, no31, no21, po41
#that means I need to not have -999 in any of those

#bats_surf_sub=
  #bats_co2sys_surf %>%
  #select(year, month, Temp, Sal1, `O2(1)`,`NO31`,`NO21`, `PO41`) %>%
  #replace_with_na_all(condition =~.x ==-999)
#bats_surf_sub=bats_surf_sub[complete.cases(bats_surf_sub),]
#Be cautious removing missing data== see online lecture and notes about potential systematic biases.

bats_surf_sub=
  bats_co2sys_surf %>%
  filter(year!=-999,month!=-999,Sal1!=-999,Temp!=-999,`O2(1)`!=-999,`NO31`!=-999,`NO21`!=-999,`PO41`!=-999)

#Now we have all of our predictors without -999 missing data

step(lm(pCO2insitu~1,data=bats_surf_sub),
     direction="forward",
     trace = 1,
     scope = ~year+month+Sal1+Temp+`O2(1)`+`NO31`+`NO21`+`PO41`)

model_AIC=lm(pCO2insitu ~ Temp + year + NO31 + Sal1 + NO21, data = bats_surf_sub)
summary(model_AIC)
check_model(model_AIC)

bats_surf_sub_pred=
  cbind(bats_surf_sub,
        predict(model_AIC, interval='confidence', level=0.95))

bats_surf_sub_pred %>% 
  ggplot()+
  geom_point(mapping=aes(x=decy,y=pCO2insitu))+
  geom_line(aes(decy,fit),color="purple")+
  geom_ribbon(aes(x=decy,ymin=lwr,ymax=upr),alpha=0.2,fill="purple")+
  theme_classic()

bats_surf_sub_pred %>% 
  ggplot()+
  geom_point(mapping=aes(x=decy,y=pCO2insitu))+
  theme_classic()
```